---
title: "03-execises"
format:
  gfm:
    number-sections: true
    toc: true
editor: source
---

## Conceptual

1. ***Describe the null hypotheses to which the p-values given in Table 3.4 correspond. Explain what conclusions you can draw based on these p-values. Your explanation should be phrased in terms of sales, TV, radio, and newspaper, rather than in terms of the coefficients of the linear model.***

Null hypotheses for each predictor each coefficient is 0. We can see in the table that we can reject the null hypotheses for **TV** and **radio** but there isn't enough evidence to reject the null hypotheses for **newspaper**.

2. ***Carefully explain the differences between the KNN classifier and KNN regression methods.***

The classifier assigns classes based on the most often class of the closest $K$ elements, on the other hand the regression estimate each value taking the mean of the closest $K$ elements.

3. ***Suppose we have a data set with five predictors to predict the starting salary after graduation (in thousands of dollars) and after using least squares we fitted the next model:***

|Variable|Coefficient|
|:-------|:----------|
|Level (High School)|$\hat{\beta}_{0} = 50$|
|$X_{1}$ = GPA|$\hat{\beta}_{1} = 20$|
|$X_{2}$ = IQ|$\hat{\beta}_{2} = 0.07$|
|$X_{3}$ = Level (College)|$\hat{\beta}_{3} = 35$|
|$X_{4}$ = Interaction between GPA and IQ|$\hat{\beta}_{4} = 0.01$|
|$X_{5}$ = Interaction between GPA and Level|$\hat{\beta}_{5} = −10$|

- **Which answer is correct, and why?***

Based on this information we can say that:

> For a fixed value of IQ and GPA, college graduates earn more, on average, than high school graduate.

As High School students earn on average $\hat{\beta}_{0} = 50$ College students earn |$\hat{\beta}_{0} + \hat{\beta}_{3} = 85$

- ***Predict the salary of a college graduate with IQ of 110 and a GPA of 4.0.***

$$
\begin{split}
\hat{Y} & = 35 + 20 (4) + 0.07 (110) + 35 + 0.01(4)(110) - 10 (4) \\
        & = 122.1
\end{split}
$$

- ***True or false: Since the coefficient for the GPA/IQ interaction term is very small, there is very little evidence of an interaction effect. Justify your answer.***

FALSE, we can not make conclusions about the significance of any tern about checking the the standard error of each term. The coefficient might small because the IQ has very high values if we contrast the GPA ones.


4. ***I collect a set of data (n = 100 observations) containing a single predictor and a quantitative response. I then fit a linear regression model to the data, as well as a separate cubic regression, i.e. $Y = \beta_{0} + \beta_{1}x + \beta_{2}x^2 + \beta_{3}x^3 + \epsilon$.***

- ***Suppose that the true relationship between X and Y is linear, i.e. $Y = \beta_{0} + \beta_{1}x + \epsilon$. Consider the training residual sum of squares (RSS) for the linear regression, and also the training RSS for the cubic regression. Would we expect one to be lower than the other, would we expect them to be the same, or is there not enough information to tell? Justify your answer.***

As the training RSS always gets lower as we increase the flexibility the cubic regression would have a lower RSS.

- ***Answer (a) using test rather than training RSS.**

The linear regression would have a lower test RSS, as it reduces de scare bias of the model.

- ***Suppose that the true relationship between X and Y is not linear, but we don’t know how far it is from linear. Consider the training RSS for the linear regression, and also the training RSS for the cubic regression. Would we expect one to be lower than the other, would we expect them to be the same, or is there not enough information to tell? Justify your answer.***

As the training RSS always gets lower as we increase the flexibility the cubic regression would have a lower RSS.

- ***Answer (c) using test rather than training RSS.**

The cubic regression would have a lower test RSS, as it reduces de scare bias of the model.


5. ***Consider the fitted values that result from performing linear regression without an intercept. In this setting, the $i$th fitted value takes the form.***

$$
\hat{y}_{i} = x_{i}\hat{\beta}
$$

***Where***

$$
\hat{\beta}= \left( \sum_{i=1}^{n}{x_{i}y_{i}}  \right) /
             \left( \sum_{i'=1}^{n}{x_{i'}^2}  \right)
$$

- ***Show that we can write***

$$
\hat{y}_{i} = \sum_{i'=1}^{n}{a_{i'}y_{i'}}
$$
I am not sure about this execise as I don't understand the difference between $i$ and $i'$.

$$
\begin{split}
\sum_{i'=1}^{n}{a_{i'}y_{i'}} & = x_{i}\hat{\beta} \\
\sum_{i'=1}^{n}{a_{i'}y_{i'}} & = x_{i}\frac{\sum_{i=1}^{n}{x_{i}y_{i}}}
                                            {\sum_{i'=1}^{n}{x_{i'}^2} } \\
\sum_{i'=1}^{n}{a_{i'}} \sum_{i'=1}^{n}{y_{i'}} & = \frac{x_{i}\sum_{i=1}^{n}{x_{i}}}
                                                         {\sum_{i'=1}^{n}{x_{i'}^2} } 
                                                    \sum_{i=1}^{n} {y_{i}} \\
\sum_{i'=1}^{n}{a_{i'}} & = \frac{x_{i}\sum_{i=1}^{n}{x_{i}}}
                                                         {\sum_{i'=1}^{n}{x_{i'}^2} }
\end{split}
$$

6. ***Using (3.4), argue that in the case of simple linear regression, the least squares line always passes through the point $(\overline{x},\overline{x})$.***

As you can see bellow the intercept it's the responsible for that property.

$$
\begin{split}
\hat{y} & = \left( \hat{\beta}_{0} \right) + \hat{\beta}_{1} \overline{x} \\
\hat{y} & = \overline{y} - \hat{\beta}_{1}\overline{x} + \hat{\beta}_{1} \overline{x} \\
\hat{y} & = \overline{y}
\end{split}
$$

## Applied

7. **This question involves the use of simple linear regression on the Auto data set.**

- ***Use the lm() function to perform a simple linear regression with mpg as the response and horsepower as the predictor. Use the summary() function to print the results. Comment on the output.***

```{r}
library(ISLR2)

AutoSimpleModel <- lm(mpg ~ horsepower, data = Auto)

summary(AutoSimpleModel)

```

As we see the regression p-value is much lower than 0.05 and we can reject the null hypotheses to conclude that there is a  **strong relationship** between the response en the predictor. The coefficient of horsepower is negative, so we know that as the predictor increase the response decrease. 

  - ***What is the predicted mpg associated with a horsepower of 98? What are the associated 95 % confidence and prediction intervals.***

```{r}
predict(AutoSimpleModel, newdata = data.frame(horsepower = 98), interval = "confidence")
predict(AutoSimpleModel, newdata = data.frame(horsepower = 98), interval = "prediction")
```

- ***Plot the response and the predictor. Use the abline() function to display the least squares regression line.***

```{r}
plot(Auto$horsepower,Auto$mpg)
abline(AutoSimpleModel)
```

- ***Use the plot() function to produce diagnostic plots of the least squares regression fit. Comment on any problems you see with the fit.***

```{r}
par(mfrow = c(2, 2))
plot(AutoSimpleModel)
```

```{r}
#| echo: false
#| include: false

par(mfrow = c(1, 1))
```


The *Residuals vs Fitted* shows that the relation is not linear and variance isn't constant.

9. ***This question involves the use of multiple linear regression on the Auto data set.***

- ***Produce a scatterplot matrix which includes all of the variables in the data set.***

```{r}
pairs(Auto)
```

- ***Compute the matrix of correlations between the variables using the function cor(). You will need to exclude the name variable, which is qualitative.***

```{r}
Auto |>
  subset(select = -name)|>
  cor()
```

- ***Use the lm() function to perform a multiple linear regression with mpg as the response and all other variables except name as the predictors. Use the summary() function to print the results. Comment on the output. For instance***

```{r}
AutoModelNoInteraction <- 
  lm(mpg ~ . -name, data = Auto)

AutoModelNoInteractionummary <- 
  summary(AutoModelNoInteraction)

AutoModelNoInteractionummary

```

  - ***Is there a relationship between the predictors and the response?***
  
  As the regression p-value is bellow 0.05 we can reject the null hypothesis and conclude that at least one of the predictors have a relation with the response.
  
  - ***Which predictors appear to have a statistically significant relationship to the response?***
  
```{r}
AutoModelNoInteractionummary |>
  coefficients() |>
  as.data.frame() |>
  subset(`Pr(>|t|)` < 0.05)
```
  
  - ***What does the coefficient for the year variable suggest?***
  
  It suggests that cars in average cars can drive 0.75 more miles per gallon every year.
  
- ***Use the plot() function to produce diagnostic plots of the linear regression fit. Comment on any problems you see with the fit. Do the residual plots suggest any unusually large outliers? Does the leverage plot identify any observations with unusually high leverage?***

  - Non-linearity of the response-predictor relationships
  - Non-constant variance
  - High-leverage points

```{r}
par(mfrow = c(2, 2))
plot(AutoModelNoInteraction)
```

```{r}
#| echo: false
#| include: false

par(mfrow = c(1, 1))
```

- ___Use the \* and : symbols to fit linear regression models with interaction effects. Do any interactions appear to be statistically significant?___


```{r}
remove_rownames <- function(DF){
  
  DF <- cbind(name = row.names(DF), DF)
  rownames(DF) <- NULL
  return(DF)
  
}


names(Auto) |>
  setdiff(c("mpg","name")) |>
  (\(x) c(x,
          combn(x, m = 2, 
                FUN = \(y) paste0(y,collapse =":"))))() |>
  paste0(collapse = " + ") |>
  paste0("mpg ~ ", predictors = _) |>
  lm(data = Auto) |>
  summary() |>
  coef() |>
  as.data.frame() |>
  remove_rownames() |>
  subset(`Pr(>|t|)` < 0.05 | name == "year")


```

  - ***Try a few different transformations of the variables, such as $\log{x}$, $\sqrt{x}$, $x^2$. Comment on your findings.***
  
  As we can see bellow we can explain 3% more of the variability by applying log to some variables.
  
```{r}
library(data.table)

apply_fun_lm <- function(FUN,DF, trans_vars, remove_vars){
  
    as.data.table(DF
    )[, (trans_vars) := lapply(.SD, FUN), .SDcols = trans_vars
    ][, !remove_vars, with = FALSE
    ][, lm(mpg ~ . , data = .SD)] |>
    summary() |>
    (\(x) data.table(adj.r.squared = x$adj.r.squared,
                     sigma = x$sigma,
                     p.value = pf(x$fstatistic["value"], 
                                  x$fstatistic["numdf"], 
                                  x$fstatistic["dendf"], 
                                  lower.tail = FALSE)))()
  
}


data.table(function_name = c("original","log", "sqrt","x^2"),
           function_list = list(\(x) x,log, sqrt, \(x) x^2)
)[, data :=  
    lapply(function_list, 
           FUN = apply_fun_lm,
           DF = Auto, 
           trans_vars = c("displacement", "horsepower", 
                          "weight", "acceleration"),
           remove_vars = "name")
][, rbindlist(data) |> cbind(function_name, end = _)]

  
```


10. ***This question should be answered using the Carseats data set.***

  - ***Fit a multiple regression model to predict Sales using Price, Urban, and US.***
  
```{r}
CarseatsModel <-
  lm(Sales~Price+Urban+US, data = Carseats)

CarseatsModelSummary <-  
  summary(CarseatsModel)

CarseatsModelSummary
```
  
  - ***Provide an interpretation of each coeﬃcient in the model. Be careful—some of the variables in the model are qualitative!***
  
```{r}

CarseatsInterationModel <-
  lm(Sales~Price*Urban*US, data = Carseats)

CarseatsInterationModelSummary <-  
  summary(CarseatsInterationModel)

CarseatsInterationModelSummary

```

  - ***Write out the model in equation form, being careful to handle the qualitative variables properly.***
  
```{r}
coef(CarseatsInterationModel) |>
  round(3) |>
  (\(x) paste0(ifelse(x < 0, " - "," + "), abs(x)," \text{ ", names(x),"}"))() |>
  sub(pattern = " \text{ (Intercept)}",replacement = "", fixed = TRUE) |>
  paste0(collapse = "") |>
  sub(pattern = "^ \\+ ", replacement = "") |>
  sub(pattern = "^ - ", replacement = "") |>
  paste0("hat{Y} = ", FUN = _)
```
  
$$
\begin{split}
\hat{Sales} & = 13.456 - 0.062 \text{ Price} - 0.652 \text{ UrbanYes} \\
            & \quad + 2.049 \text{ USYes} + 0.011 \text{ Price:UrbanYes} \\
            & \quad - 0.002 \text{ Price:USYes} - 1.122 \text{ UrbanYes:USYes} \\
            & \quad + 0.001 \text{ Price:UrbanYes:USYes}
\end{split}
$$


  - ***For which of the predictors can you reject the null hypothesis H0 : βj = 0?***
  
```{r}

coef(CarseatsInterationModelSummary) |>
  as.data.frame() |>
  (\(DF) DF[DF$`Pr(>|t|)` < 0.05,])()

```
  
  - ***On the basis of your response to the previous question, ﬁt a smaller model that only uses the predictors for which there is evidence of association with the outcome.***  

```{r}
CarseatsPriceModel <-
  lm(Sales~Price, data = Carseats)

CarseatsPriceModelSummary <-
  summary(CarseatsPriceModel)

CarseatsPriceModelSummary

```

  - ***How well do the models in (a) and (e) ﬁt the data?***
  
**Model a** fits better to the data with `r round(CarseatsModelSummary$adj.r.squared, 2)` against `r round(CarseatsPriceModelSummary$adj.r.squared, 2)` of **model e**.

 - ***Using the model from (e), obtain 95 % conﬁdence intervals for the coeﬃcient(s).***
 
```{r}
confint(CarseatsPriceModel, level = 0.95)
```


  - ***Is there evidence of outliers or high leverage observations in the model from (e)?***
  
```{r}
par(mfrow = c(2,2))
plot(CarseatsPriceModel)
par(mfrow = c(1,1))
```

There is a leverage point.

11. ***In this problem we will investigate the t-statistic for the null hypothesis H0 : β = 0 in simple linear regression without an intercept. To begin, we generate a predictor x and a response y as follows.***

```{r}
set.seed(1)
x<-rnorm(100)
y<-2*x+rnorm(100)
```


